Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-03-21 ribose-paper.RMD
updated discussion

The results from the strength tests do indicate that the participants might have been fatigued during the intervention, due to exercising every day for 12 days straight, with strength testing in-between. In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing compared to placebo. However, this was generally not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions. Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, since glucose is the preferred energy source during strenuous activity [@mul_exercise_2015]. Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance during strength testing. This could also explain the differences between treatments in content of the stress responsive c-Myc [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc. If this was the case, it does seem inconclusive that strength did not differ between treatments at earlier or later time points. In addition, there were no differences in training volume per session between treatments, therefore, the mechanical loading should have been equal. Another potential explanation to this observation might be …. (bifasisk restitusjon?)<!-- investigate this -->. Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design. Initially, we met our minimum limit of 16 participants however, three dropped out during the intervention, leaving us slightly underpowered according to the a priori calculation.

>>>>

As in the biological data, there was generally no difference in change of strength as an index between the treatments. Generally, skeletal muscle strength decreased from baseline to post fifth and sixth RT session, with a significantly less reduction in glucose compared to placebo at post fifth session measurement. This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity was quite high. Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@RAASTAD2000]. In their study on recovery of skeletal muscle contractility after strength exercise, Raastad and Hallén (2000) found a rapid recovery during the initial 11hrs post-exercise, followed by either a leveling off or a drop until 22hrs post-exercise, returning to baseline values at 33hrs post-exercise. Herein, inflammation and phagocytic activity was proposed to be involved in th performance drop between 11-22hrs [@RAASTAD]. Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post sixth RT session and a drop at 23hrs. Moreover, the strength measurements during the intervention and after five RT sessions was also conducted approximately 23hrs after RT. Thus, it is possible that these strength measurements also was affected by the biphasic recovery pattern. However, this does not explain the difference between treatments at post fifth RT session. A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increased performance in RT with glucose compared to RT with placebo. This may also, at least partly, explain significantly higher content of the stress responsive c-Myc in RT with placebo compared to glucose [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. However, training volume data do reveal that the total training session volume was equal on pairwise consecutive days, i.e. no difference between day 1-2, day 3-4 and so on. Thus, there were no differences in mechanical loading to induce higher stress in RT with placebo compared to RT with glucose. Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in a lower metabolic stress, compared to RT with placebo [..] <!-- REFS --> Unfortunately, we did not conduct measurements of inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio, thus, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however possible. Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design. Initially, we met our minimum limit of 16 participants however, three dropped out during the intervention, leaving us slightly underpowered according to the a priori calculation.
